---
title: "p8105_hw3_mnt2130"
author: "Mindy Tran"
date: "2022-10-13"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


###Problem 1

This code chunk will load  and read the data from the p8105 datasets. 

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Each row contains data about individual items in an instacart order. Variables include id numbers that were used to identify each order, the user who made the order, and the products in the order and the the sequential order in which each product was added to the cart. There are variables that tell us more about each order such as when the order was made (day and hour of day), and number of days since prior order. There are also other variables that describe each product- giving us the product name, department,the aisle where its found, and whether this product is a reorder.  From this data, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Here is a table that tells us the the total number of items ordered from each aisle. Out of the 134 aisles, fresh vegetables are the most ordered item.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Below is a plot that shows the number of items ordered in each aisle in ascending order. 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

From this plot, butter, oils/ vinegars are some of the least ordered items on Instacart. 

The code below  generates a table showing the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits` with the number of times each item was ordered. The most ordered vegetable is organic baby spinuach, the most ordered dog food care item is snack sticks dog treats and the most ordered baking ingredient is light brown sugar. 

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```
The most ordered vegetable is organic baby spinuach, the most ordered dog food care item is snack sticks dog treats and the most ordered baking ingredient is light brown sugar. 

Here is a table that shows the mean hour of the day that Pink Lady Apples and Coffee Ice Cream were ordered on each day of the week. This table has been formatted for human readers in an untidy format.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

###Problem 2

